|

### 2021-05-29 article/article.Rmd
revise methods section, add results

---
title:  |
        | The Effect of Legislature Size on Public Spending:
        | A Meta-Analysis
author: 
  - Huzeyfe Alptekin[^huzeyfe]
  - Danilo Freire[^danilo]
  - Umberto Mignozzetti[^umberto]
  - Catarina Roman[^catarina]
date: \today
thanks: "The authors thank Guilherme Duarte, Robert Myles McDonnell, and David Skarbek for their constructive feedback. We also thank Cedric Antunes, Luis Castro, Giovanna França, Julia Oriente, and Lucas Mingardi for their excellent research assistance. Replication materials are available at <https://github.com/danilofreire/distributive-politics-meta-analysis>. We kindly acknowledge funding from the São Paulo State Science Foundation (FAPESP grant number 2018/00646-1)."
abstract: "In a seminal article, @weingast1981political argue that there is a positive relationship between legislature size and inefficiency in public expenditures. Their proposition is currently known as the \"law of $1/n$\" and has been widely debated in political science and public administration. However, recent studies have questioned the validity of the theory. In this letter, we estimate the first meta-analysis of the relationship between the number of legislators and public spending. Based on a sample of 29 articles, we find no robust evidence for the effect of legislature size on government budgets. Yet the aggregate results show significant heterogeneity. While earlier studies provide moderate support for the \"law of $1/n$\", papers using causal inference methods consistently find a negative relationship between seats and spending. The available evidence also indicates that proportional representation and mixed voting systems are no more likely to overspend than majoritarian ones."
abstractspacing: double
keywords: distributive politics; law of $1/n$; legislature size; meta-analysis; public spending
jelcodes: H21; H23; H50; H61
fontsize: 11pt
margin: 2cm
urlcolor: darkblue
linkcolor: Mahogany
citecolor: Mahogany
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    keep_tex: no
    toc: no
    toc_depth: 3
    template: article-template.latex
    md_extensions: +raw_attribute
---

@@

Since many empirical tests of the "law of $1/n$" have produced conflicting
results, scholars have expanded this research agenda and investigated how
institutional factors condition the original formulation of the theory. For
instance, authors such as @crowley2019law and @pecorino2018supermajority
accurately point out that collective action problems have been overlooked in
the literature, and recent findings indicate that bicameralism
[@maldonado2013legislatures], intergovernmental competition
[@crowley2015local], redistricting [@lee2018court], and party ideology
[@bjedov2014impact] strongly influence the relationship between seats and
spending. Moreover, the literature has increasingly applied causal inference
designs to estimate the effect of the "law of $1/n$", and in contrast to
previous studies using panel data, regression discontinuity designs generally
indicate that more legislators decrease public expenditures
[@debenedetto2018effect; @hohmann2017effect; @lewis2019legislature]. In this
respect, scholars have long been aware of the theoretical and empirical
limitations of the "law of $1/n$", and the proliferation of new studies reflect
a conscious attempt to assess the robustness of the theory [@crowley2019law].

In this letter, we conduct the first meta-analysis that tests the generality of
the "law of $1/n$". We select `r length(unique(dat$id))` articles that
use quantitative methods to evaluate the impact of legislature size over
government spending across several dimensions. Our study sample mirrors the
diversity of the literature. Out of the `r dim(dat)[1]` coefficients
included in our main analysis, `r as.numeric(round(prop.table(table(dat$scoefnd))*100, 1)[which(names(prop.table(table(dat$scoefnd)))=='Positive Significant')])`\% of
them are positive and statistically significant, `r as.numeric(round(prop.table(table(dat$scoefnd))*100, 1)[which(names(prop.table(table(dat$scoefnd)))=='Positive Insignificant')])`\%
are positive and statistically insignificant, `r as.numeric(round(prop.table(table(dat$scoefnd))*100, 1)[which(names(prop.table(table(dat$scoefnd)))=='Negative Insignificant')])`\%
are negative and statistically insignificant, and `r as.numeric(round(prop.table(table(dat$scoefnd))*100, 1)[which(names(prop.table(table(dat$scoefnd)))=='Negative Significant')])`\% are
negative and statistically significant. Meta-analysis provides a rigorous
approach to combine heterogeneous outcomes into a single estimation, and it
allows scholars to gain valuable insights from the aggregated data
[@cooper2019handbook; @hedges1985statistical]. Meta-analysis can also identify
potential sources of study variability, enabling researchers to assess threats
to external validity and direct future efforts into more promising areas of
academic inquiry [@doucouliagos2008democracy]. Research synthesis methods have
been successfully used in medicine and psychology since the 1970s
[@glass2015meta], and our work contributes to the burgeoning literature that
applies meta-analytic methods to understand challenging questions in political
science [@costa2017responsive; @doucouliagos2008democracy; @green2013field;
@lau2007effects; @schwarz2020supporting].

We run binomial tests and meta-regressions to further evaluate the relationship
between legislature size and government expenditures. We use the binomial tests
to asses whether the proportion of positive and negative coefficients in our
sample are statistically different from each other. In the meta-regressions, we
measure the effect of five moderators that capture different sources of
heterogeneity in the literature. As our sample includes both published and
unpublished articles, we also construct a series of funnel plots to test for
publication bias in existing research on the "law of $1/n$"
[@easterbrook1991publication; @gerber2001testing].

Aggregate results indicate that legislature size has no significant impact on
public spending. All of our main meta-analysis estimates show that the overall
effect is not statistically different from zero, thus confirming the
conflicting findings reported by the literature. However, when we look only at
articles that employ regression discontinuity designs (RDDs), we see that all
three papers included in our sample suggest that a higher number of legislators
leads to lower public spending [@debenedetto2018effect; @hohmann2017effect;
@lewis2019legislature]. In this regard, it is possible that methodological
choices explain part of the disparity observed in the literature, as studies
which use causal inference methods consistently point to the same direction.
One limitation of this finding is that these studies only test the impact of
lower house size on the natural logarithm of expenditure per capita, and it
remains unclear whether the results hold with other variables.

The meta-regressions provide additional evidence that our study sample is
highly heterogeneous and that effect sizes differ substantially according to
study specifications. When using our extended sample of `r dim(fulldat)[1]`
coefficients, we find that unicameralism is associated with higher public
spending, as predicted by the "law of $1/n$". Since most unicameral
legislatures are located in local governments (municipalities or districts),
this result supports the theory strictly in its first formulation. Moreover,
our meta-regressions indicate that larger upper chambers spend more in terms of
per capita expenditure than lower chambers, a result that also appears in the
binomial tests [@gilligan1995deviations; @gilligan2001fiscal]. Non-majoritarian
voting systems decrease government spending as a percentage of GDP, but they do
not have a relevant impact on the other measures of public expenditure,
following the idea that the $1/n$ effect grows weaker as the empirical cases
distance from the original definition of the law. Finally, meta-regression
results confirm that regression discontinuity models reduce public spending
estimates.

@@

In the second round, we did not require articles to cite @weingast1981political
and used a keyword-based query on Google Scholar to broaden the scope of the
first search. The search string contained terms strongly associated with the
literature on the "law of $1/n$" and was as follows: `("upper chamber size" OR
"lower chamber size" OR "council size" OR "parliament size" OR "legislature
size" OR "number of legislators" OR "legislative size") AND ("spending" OR
"expenditure" OR "government size")`. We again restricted the search to
articles written in English which employed quantitative methods. This search
added two new results to our sample (**CITE**), but neither of them included
variables beyond the six measures we had previously identified. In the third
search round, we looked into the personal webpages of every author whom we had
already included in our sample. The purpose of this manual search was to assess
whether there was any working paper or unpublished manuscript that we had
missed in the two former queries. All papers we found in the last search had
already been included in our sample. Combined, the three searches produced a
dataset of `r length(unique(dat$id))` studies as of the 10\textsuperscript{th}
of March 2021. Table \ref{tab:papers} contains the full list of articles we analyse in
this paper.

\footnotesize
\begin{longtable}{>{\raggedright\arraybackslash}p{3.9cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2.3cm}>{\centering\arraybackslash}p{1cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{1cm}}
\caption{Papers included in the meta-analysis, ordered by year of appearance}\\
\toprule
\raggedright Author(s) & \centering Journal & \centering Country & \centering Dependent Variable & \centering Method & \centering Institutional Design & \centering Electoral System
\tabularnewline
\midrule
\endhead 
```{r papers, echo = FALSE, results = 'asis', cache = T}
build_nams <- function(x) {
  x_ <- unlist(strsplit(as.character(x), ' & ', fixed = T, useBytes = T))
  if(length(x_) == 1) return(x_)
  else {
    aux <- x_[1]
      for (i in 2:length(x_))
        if (i == length(x_)){
          aux <- paste(aux, ' \\& \\\\ ', x_[i], sep = '')
        } else {
          aux <- paste(aux, ',\\\\ ', x_[i], sep = '')
        }
  }
  return(paste('\\begin{tabular}[t]{@{}l@{}}', aux,'\\end{tabular}', 
                   sep = ''))
}
aux <- dat %>%
  select(authoryear, journalCode, locationISO, depvar2, method, instdesign, elecsys3, year) %>%
  unique() %>%
  arrange(year)
for (i in 1:dim(aux)[1]) {
  cat(paste(build_nams(aux$authoryear[i]), ' & ', aux$journalCode[i], ' & ',
      aux$locationISO[i], ' & ', aux$depvar2[i], ' & ', 
      aux$method[i], ' & ', aux$instdesign[i], ' & ', 
      aux$elecsys3[i], ' \\\\ [0.5ex]', sep = ''), '\n')
}
```
\bottomrule
\label{tab:papers}
\begin{minipage}{\textwidth}
\renewcommand{\footnoterule}{}
\vspace{-0.5cm}
\footnotetext{\textbf{Note:} Journals: Unpub$=$Unpublished, JPE$=$Journal of
Political Economy, EJPE$=$European Journal of Political Economy, PC$=$Public
Choice, JPubE$=$Journal of Public Economics, JPriE$=$Journal of Private
Enterprise, APSR$=$American Political Science Review, SEJ$=$Southern Economic
Journal, UAR$=$Urban Affairs Review, SCID$=$Studies in Comparative
International Development, SSQ$=$Social Science Quarterly, SPPQ$=$State
Politics and Policy Quarterly, CPS$=$Comparative Political Studies,
RivPE$=$Rivista di Politica Economica, E\&P$=$Economics and Politics,
NTJ$=$National Tax Journal. Country codes follow the ISO 3166-1 alpha-3
international standard. Dependent variable: ExpPC$=$Per capita expenditure,
logExpPC$=$Natural logarithm of per capita expenditure, PCTGDP$=$Expenditure as
a percentage of GDP. Method: OLS$=$Ordinary least squares, IV$=$Instrumental
variables, Panel$=$Panel data/fixed effects, RDD$=$Regression discontinuity
design. Electoral system: M$=$Majoritarian, NM$=$Non-majoritarian (mixed or
proportional representation).} 
\end{minipage} 
\end{longtable}

\normalsize

Our study sample reflects the development of the literature. Although the "law
of $1/n$" was first formulated in 1981, the empirical assessment of the theory
only started a few years later, as dates of publishing range from `r min(dat$year)`
to `r max(dat$year)`. Most studies focus on the United States (`r as.numeric(table(unique(dat[,c('id','locationISO')])$locationISO)[which(names(table(unique(dat[,c('id','locationISO')])$locationISO))=='USA')])`), but our sample also contains
papers on Australia (`r as.numeric(table(unique(dat[,c('id','locationISO')])$locationISO)[which(names(table(unique(dat[,c('id','locationISO')])$locationISO))=='AUS')])`), Brazil (`r as.numeric(table(unique(dat[,c('id','locationISO')])$locationISO)[which(names(table(unique(dat[,c('id','locationISO')])$locationISO))=='BRA')])`), Germany (`r as.numeric(table(unique(dat[,c('id','locationISO')])$locationISO)[which(names(table(unique(dat[,c('id','locationISO')])$locationISO))=='DEU')])`), Indonesia (`r as.numeric(table(unique(dat[,c('id','locationISO')])$locationISO)[which(names(table(unique(dat[,c('id','locationISO')])$locationISO))=='IDN')])`), Italy (`r as.numeric(table(unique(dat[,c('id','locationISO')])$locationISO)[which(names(table(unique(dat[,c('id','locationISO')])$locationISO))=='ITA')])`), and
Switzerland (`r as.numeric(table(unique(dat[,c('id','locationISO')])$locationISO)[which(names(table(unique(dat[,c('id','locationISO')])$locationISO))=='CHE')])`). Eight articles use cross-national data and analyse from 2 to 110
countries. Early studies used OLS and panel data methods to estimate the
results, while studies from 2012 onward have also applied causal inference designs,
such as instrumental variables and regression discontinuity models.

@@

>>>>

---
title:  |
        | The Effect of Legislature Size on Public Spending:
        | A Meta-Analysis
author: 
  - Huzeyfe Alptekin[^huzeyfe]
  - Danilo Freire[^danilo]
  - Umberto Mignozzetti[^umberto]
  - Catarina Roman[^catarina]
date: \today
thanks: "The authors thank Guilherme Duarte, Robert Myles McDonnell, and David Skarbek for their constructive feedback. We also thank Cedric Antunes, Luis Castro, Giovanna França, Julia Oriente, and Lucas Mingardi for their excellent research assistance. Replication materials are available at <https://github.com/danilofreire/legislature-size-meta-analysis>. We kindly acknowledge funding from the São Paulo State Science Foundation (FAPESP grant number 2018/00646-1)."
abstract: "In a seminal article, @weingast1981political argue that there is a positive relationship between legislature size and inefficiency in public expenditures. Their proposition is currently known as the \"law of $1/n$\" and has been widely debated in political science and public administration. However, recent studies have questioned the validity of the theory. In this letter, we estimate the first meta-analysis of the relationship between the number of legislators and public spending. Based on a sample of 29 articles, we find no robust evidence for the effect of legislature size on government budgets. Yet the aggregate results show significant heterogeneity. While earlier studies provide moderate support for the \"law of $1/n$\", papers using causal inference methods consistently find a negative relationship between seats and spending. The available evidence also indicates that proportional representation and mixed voting systems are no more likely to overspend than majoritarian ones."
abstractspacing: double
keywords: distributive politics; law of $1/n$; legislature size; meta-analysis; public spending
jelcodes: H21; H23; H50; H61
fontsize: 11pt
margin: 2cm
urlcolor: darkblue
linkcolor: Mahogany
citecolor: Mahogany
spacing: double
papersize: a4paper
bibliography: references.bib
biblio-style: apalike
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    number_sections: yes
    keep_tex: no
    toc: no
    toc_depth: 3
    template: article-template.latex
    md_extensions: +raw_attribute
---

@@

Since many empirical tests of the "law of $1/n$" have produced conflicting
results, scholars have expanded this research agenda and closely investigated
how institutional factors condition the original formulation of the theory. For
instance, authors such as @crowley2019law and @pecorino2018supermajority
accurately point out that collective action problems have been overlooked in
the literature, and recent findings indicate that bicameralism
[@maldonado2013legislatures], intergovernmental competition
[@crowley2015local], redistricting [@lee2018court], and party ideology
[@bjedov2014impact] strongly influence the relationship between seats and
spending. Moreover, the literature has increasingly applied causal inference
designs to estimate the effect of the "law of $1/n$", and in contrast to
previous studies using panel data, regression discontinuity designs generally
indicate that more legislators decrease public expenditures
[@debenedetto2018effect; @hohmann2017effect; @lewis2019legislature]. In this
respect, scholars have long been aware of the theoretical and empirical
limitations of the "law of $1/n$", and the proliferation of new studies reflect
a conscious attempt to assess the robustness of the theory [@crowley2019law].

In this letter, we conduct the first meta-analysis that tests the generality of
the "law of $1/n$". We select `r length(unique(dat$id))` articles that
use quantitative methods to evaluate the impact of legislature size over
government spending across several dimensions. Our study sample mirrors the
diversity of the literature. Out of the `r dim(dat)[1]` coefficients
included in our main analysis, `r as.numeric(round(prop.table(table(dat$scoefnd))*100, 1)[which(names(prop.table(table(dat$scoefnd)))=='Positive Significant')])`\% of
them are positive and statistically significant, `r as.numeric(round(prop.table(table(dat$scoefnd))*100, 1)[which(names(prop.table(table(dat$scoefnd)))=='Positive Insignificant')])`\%
are positive and statistically insignificant, `r as.numeric(round(prop.table(table(dat$scoefnd))*100, 1)[which(names(prop.table(table(dat$scoefnd)))=='Negative Insignificant')])`\%
are negative and statistically insignificant, and `r as.numeric(round(prop.table(table(dat$scoefnd))*100, 1)[which(names(prop.table(table(dat$scoefnd)))=='Negative Significant')])`\% are
negative and statistically significant. Given the volume and the disparity of
the studies, we employ meta-analysis to summarise the results. Meta-analysis
provides a rigorous approach to combine heterogeneous outcomes into a single
estimation, and it allows scholars to gain valuable insights from the
aggregated data [@cooper2019handbook; @hedges1985statistical]. Meta-analysis
can also identify potential sources of study variability, enabling researchers
to assess threats to external validity and direct future efforts into more
promising areas of academic inquiry [@doucouliagos2008democracy]. Research
synthesis methods have been successfully applied in medicine and psychology
since the 1970s [@glass2015meta], and our work contributes to the burgeoning
literature that uses meta-analytic methods to understand challenging questions
in political science [@costa2017responsive; @doucouliagos2008democracy;
@green2013field; @lau2007effects; @schwarz2020supporting].

We run binomial tests and meta-regressions to further evaluate the relationship
between legislature size and government expenditures. We use binomial tests
to assess whether the proportions of positive and negative coefficients in our
sample are statistically different from each other. In the meta-regressions, we
measure the effect of five moderators that capture different sources of
heterogeneity in the literature. As our sample includes both published and
unpublished articles, we also construct a series of funnel plots to test for
publication bias in existing research on the "law of $1/n$"
[@easterbrook1991publication; @gerber2001testing].

Aggregate results indicate that legislature size has no significant impact on
public spending. All of our main meta-analysis estimates show that the overall
effect is not statistically different from zero, thus confirming the
conflicting findings reported by the literature. However, when we look only at
articles that employ regression discontinuity designs, we see that all three
papers included in our sample suggest that a higher number of legislators leads
to lower public spending [@debenedetto2018effect; @hohmann2017effect;
@lewis2019legislature]. In this regard, it is possible that methodological
choices partially explain the divergent results observed in the literature, as
studies that use causal inference methods consistently point to the same
direction. One limitation of this finding is that these studies only test the
impact of lower house size on the natural logarithm of expenditure per capita,
thus it remains unclear whether the results hold with other variables.

The meta-regressions provide additional evidence that our study sample is
highly heterogeneous and that effect sizes differ substantially according to
study specifications. When using our extended sample of `r dim(fulldat)[1]`
coefficients, we find that unicameralism is associated with higher public
spending, as predicted by the "law of $1/n$". Since most unicameral
legislatures are located in local governments (municipalities or districts),
this result supports the theory strictly in its first formulation. Moreover,
our meta-regressions indicate that larger upper chambers spend more in terms of
per capita expenditure than lower chambers, a result that also appears in the
binomial tests. Non-majoritarian voting systems decrease government spending as
a percentage of GDP, but they do not have a relevant impact on the other
measures of public expenditure, following the idea that the $1/n$ effect grows
weaker as the empirical cases distance from the original definition of the law.
Finally, meta-regression results confirm that regression discontinuity models
reduce public spending estimates.

@@

In the second round, we did not require articles to cite @weingast1981political
and used a keyword-based query on Google Scholar to broaden the scope of the
first search. The search string contained terms strongly associated with the
literature on the "law of $1/n$" and was as follows: `("upper chamber size" OR
"lower chamber size" OR "council size" OR "parliament size" OR "legislature
size" OR "number of legislators" OR "legislative size") AND ("spending" OR
"expenditure" OR "government size")`. We again restricted the search to
articles written in English which employed quantitative methods. This search
added two new results to our sample [@coate2011government;
@debenedetto2018effect], but neither of them included variables beyond the six
measures we had previously identified. In the third search round, we looked
into the personal webpages of every author whom we had already included in our
sample. The purpose of this manual search was to assess whether there was any
working paper or unpublished manuscript that we had missed in the two former
queries. All papers we found in the last search had already been included in
our sample. Combined, the three searches produced a dataset of 
`r length(unique(dat$id))` studies as of the 10\textsuperscript{th} of March 2021.
Table \ref{tab:papers} contains the full list of articles we analyse in this
paper.

\footnotesize
\begin{longtable}{>{\raggedright\arraybackslash}p{3.9cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{2.3cm}>{\centering\arraybackslash}p{1cm}>{\centering\arraybackslash}p{2cm}>{\centering\arraybackslash}p{1.1cm}}
\caption{Papers included in the meta-analysis, ordered by year of appearance}\\
\toprule
\raggedright Author(s) & \centering Journal & \centering Country & \centering Dependent Variable & \centering Method & \centering Institutional Design & \centering Electoral System
\tabularnewline
\midrule
\endhead 
```{r papers, echo = FALSE, results = 'asis', cache = T}
build_nams <- function(x) {
  x_ <- unlist(strsplit(as.character(x), ' & ', fixed = T, useBytes = T))
  if(length(x_) == 1) return(x_)
  else {
    aux <- x_[1]
      for (i in 2:length(x_))
        if (i == length(x_)){
          aux <- paste(aux, ' \\& \\\\ ', x_[i], sep = '')
        } else {
          aux <- paste(aux, ',\\\\ ', x_[i], sep = '')
        }
  }
  return(paste('\\begin{tabular}[t]{@{}l@{}}', aux,'\\end{tabular}', 
                   sep = ''))
}
aux <- dat %>%
  select(authoryear, journalCode, locationISO, depvar2, method, instdesign, elecsys3, year) %>%
  unique() %>%
  arrange(year)
for (i in 1:dim(aux)[1]) {
  cat(paste(build_nams(aux$authoryear[i]), ' & ', aux$journalCode[i], ' & ',
      aux$locationISO[i], ' & ', aux$depvar2[i], ' & ', 
      aux$method[i], ' & ', aux$instdesign[i], ' & ', 
      aux$elecsys3[i], ' \\\\ [0.5ex]', sep = ''), '\n')
}
```
\bottomrule
\label{tab:papers}
\begin{minipage}{\textwidth}
\renewcommand{\footnoterule}{}
\vspace{-0.5cm}
\footnotetext{\textbf{Journal:} Unpub$=$Unpublished, JPE$=$Journal of Political
Economy, EJPE$=$European Journal of Political Economy, PC$=$Public Choice,
JPubE$=$Journal of Public Economics, JPriE$=$Journal of Private Enterprise,
APSR$=$American Political Science Review, SEJ$=$Southern Economic Journal,
UAR$=$Urban Affairs Review, SCID$=$Studies in Comparative International
Development, SSQ$=$Social Science Quarterly, SPPQ$=$State Politics and Policy
Quarterly, CPS$=$Comparative Political Studies, RivPE$=$Rivista di Politica
Economica, E\&P$=$Economics and Politics, NTJ$=$National Tax Journal.}
\footnotetext{\textbf{Country:} Country codes follow the ISO 3166-1 alpha-3
international standard.} 
\footnotetext{\textbf{Dependent Variable:} ExpPC$=$Per capita expenditure,
logExpPC$=$Natural logarithm of per capita expenditure, PCTGDP$=$Expenditure as
a percentage of GDP.}
\footnotetext{\textbf{Method:} OLS$=$Ordinary least squares, IV$=$Instrumental
variables, Panel$=$Panel data/fixed effects, RDD$=$Regression discontinuity
design.}
\footnotetext{\textbf{Electoral System:} M$=$Majoritarian, NM$=$Non-majoritarian (mixed or proportional
representation).} 
\end{minipage} 
\end{longtable}
\normalsize

Our study sample reflects the development of the literature. Although the "law
of $1/n$" was first formulated in 1981, the empirical assessment of the theory
only started a few years later, as dates of publishing range from `r min(dat$year)`
to `r max(dat$year)`. Most studies focus on the United States (`r as.numeric(table(unique(dat[,c('id','locationISO')])$locationISO)[which(names(table(unique(dat[,c('id','locationISO')])$locationISO))=='USA')])`), but our sample also contains
papers on Australia (`r as.numeric(table(unique(dat[,c('id','locationISO')])$locationISO)[which(names(table(unique(dat[,c('id','locationISO')])$locationISO))=='AUS')])`), Brazil (`r as.numeric(table(unique(dat[,c('id','locationISO')])$locationISO)[which(names(table(unique(dat[,c('id','locationISO')])$locationISO))=='BRA')])`), Germany (`r as.numeric(table(unique(dat[,c('id','locationISO')])$locationISO)[which(names(table(unique(dat[,c('id','locationISO')])$locationISO))=='DEU')])`), Indonesia (`r as.numeric(table(unique(dat[,c('id','locationISO')])$locationISO)[which(names(table(unique(dat[,c('id','locationISO')])$locationISO))=='IDN')])`), Italy (`r as.numeric(table(unique(dat[,c('id','locationISO')])$locationISO)[which(names(table(unique(dat[,c('id','locationISO')])$locationISO))=='ITA')])`), and
Switzerland (`r as.numeric(table(unique(dat[,c('id','locationISO')])$locationISO)[which(names(table(unique(dat[,c('id','locationISO')])$locationISO))=='CHE')])`). Eight articles use cross-national data and analyse from 2 to 110
countries. Early studies used OLS and panel data methods to estimate the
results, while studies from 2012 onward have also applied causal inference
designs such as instrumental variables and regression discontinuity models to
test the relationship between house size and public spending.

@@

We use Hedges' $g$ to calculate effect sizes in our meta-analysis
[@hedges1981distribution]. While there are other methods to standardise
coefficients in meta-analytic studies, Hedges' $g$ corrects for upward bias in
small sample sizes and is considered more robust than measures such as Cohen's
$d$ [@lakens2013calculating]. We estimate the Standardised Mean Difference
(SMD), which represents the effect size in each study relative to the
variability observed in that study, by extracting the coefficients and the
standard errors from all articles included in our sample and converting them to
Hedges' $g$. In cases where authors did not report the standard errors for
their estimates, we computed them using the t-statistic presented in the
original tables.

# Results

The "law of $1/n$" states that more legislators increase government
expenditure. In this paper, we employ three methods to test the empirical
validity of that relationship. First, we run binomial z-tests to assess the
frequency of positive and negative coefficients in our sample. Then, we fit 9
multilevel meta-analyses using the `meta` [@balduzzi2019perform] and the
`dmetar` [@dmetar2019] packages for the `R` statistical language [@rstats2019].
Lastly, we run two sets of meta-regressions to measure the effects of a group
of moderators on the effect sizes of our study sample. To recapitulate, our
independent variables of interest are lower chamber size, the natural logarithm
of lower chamber size, and upper chamber size. The dependent variables are
public expenditure per capita, the natural logarithm of public expenditure per
capita, and government expenditure as a percentage of GDP. Since the outcomes
have different scales, we treat them separately in our models.

```{r, cache=TRUE, echo = FALSE}
aux <- filter(dat, indepvar2 == "N")
binN <- binom.test(table(aux$scoef)[2], sum(table(aux$scoef)), p = 0.5)
aux <- filter(dat, indepvar2=='K')
binK <- binom.test(table(aux$scoef)[2], sum(table(aux$scoef)), p=0.5)
aux <- filter(dat, indepvar2 == "logN")
binlogN <- binom.test(table(aux$scoef)[2], sum(table(aux$scoef)), p = 0.5)
```

## Binomial Z-Tests
\label{sub:binomial}

The binomial z-test evaluates whether the majority of the coefficients of our
independent variables are positive or negative[^binomial]. If increasing the
number of legislators leads to higher public expending, we should expect most
coefficients to be positive. Conversely, if the "reverse law of $1/n" is true,
our results should indicate that $p < 0.5$, that is, the proportion of positive
coefficients is lower than 0.5. The null hypothesis here states that the sign
of the coefficient is equally likely to be positive or negative.

[^binomial]: Since coefficients are either positive or negative, each of them
can be considered a Bernoulli trial. Thus, the aggregate results for all papers
follow a Binomial distribution where parameters $n$ equals the number of papers
and $p$ is the chance of a positive coefficient. The "law of $1/n$" can be
reformulated as the chance of $p > 0.5$, which facilitates the testing of the
theory.

We start with lower house size. Our results indicate that there is no
correlation between the number of legislators in the lower house and public
expenditure (successes = `r as.numeric(binN$statistic)`, trials = `r as.numeric(binN$parameter)`, p~success~ = `r round(as.numeric(binN$estimate), 3)`, 95\% CI = [`r round(as.numeric(binN$conf.int[1]), 3)`; `r round(as.numeric(binN$conf.int[2]), 3)`], $p$-value = `r round(as.numeric(binN$p.value), 3)`). Note that the "law of $1/n$" suggests that there is a positive
association between both. The binomial test for the natural logarithm of lower
chamber size also shows a
non-statistically significant result (successes = `r as.numeric(binlogN$statistic)`, trials = `r as.numeric(binlogN$parameter)`, p~success~ = `r round(as.numeric(binlogN$estimate), 3)`, 95\% CI = [`r round(as.numeric(binlogN$conf.int[1]), 3)`; `r round(as.numeric(binlogN$conf.int[2]), 3)`],
$p$-value = `r round(as.numeric(binlogN$p.value), 3)`). In contrast, we find a
positive result for the number of legislators in the upper house, which is
in line with the mainstream literature (successes = `r as.numeric(binK$statistic)`, trials = `r as.numeric(binK$parameter)`, p~success~ = `r round(as.numeric(binK$estimate), 3)`, 95\% CI = [`r round(as.numeric(binK$conf.int[1]), 3)`; `r round(as.numeric(binK$conf.int[2]), 3)`],
$p$-value = `r round(as.numeric(binK$p.value), 3)`).

## Meta-Analysis
\label{sub:Meta-Analysis}

```{r, echo = FALSE, message = FALSE, warning = FALSE, cache=TRUE}
mod <- list()
# Pooling effects analysis -- ExpPC x N
aux <- dat %>%
  filter(indepvar2 == 'N',
         depvar2 == 'ExpPC')

mod[[1]] <- metagen(coef, SE, data=aux,
          studlab=paste(authoryear),
          comb.fixed = FALSE,
          comb.random = TRUE,
          method.tau = "REML",
          hakn = TRUE,
          prediction=TRUE,
          sm="SMD")

# Pooling effects analysis -- ExpPC x K
aux <- dat %>%
  filter(indepvar2 == 'K',
         depvar2 == 'ExpPC')

mod[[2]] <- metagen(coef, SE, data=aux,
          studlab=paste(authoryear),
          comb.fixed = FALSE,
          comb.random = TRUE,
          method.tau = "REML",
          hakn = TRUE,
          prediction=TRUE,
          sm="SMD")
# Pooling effects analysis -- logExpPC x N
aux <- dat %>%
  filter(indepvar2 == 'N',
         depvar2 == 'logExpPC')

mod[[3]] <- metagen(
  coef, SE, data=aux,
  studlab=paste(authoryear),
  comb.fixed = FALSE,
  comb.random = TRUE,
  method.tau = "REML",
  hakn = TRUE,
  prediction = TRUE,
  sm="SMD"
  )

# Pooling effects analysis -- logExpPC x logN
aux <- dat %>%
  filter(indepvar2 == 'logN',
         depvar2 == 'logExpPC')

mod[[4]] <- metagen(coef, SE, data=aux,
          studlab=paste(authoryear),
          comb.fixed = FALSE,
          comb.random = TRUE,
          method.tau = "REML",
          hakn = TRUE,
          prediction=TRUE,
          sm="SMD")

# Pooling effects analysis -- PCTGDP x N
aux <- dat %>%
  filter(indepvar2 == 'N',
         depvar2 == 'PCTGDP')

mod[[5]] <- metagen(coef, SE, data=aux,
          studlab=paste(authoryear),
          comb.fixed = FALSE,
          comb.random = TRUE,
          method.tau = "REML",
          hakn = TRUE,
          prediction=TRUE,
          sm="SMD")

# Pooling effects analysis -- PCTGDP x logN
aux <- dat %>%
  filter(indepvar2 == 'logN',
         depvar2 == 'PCTGDP')

mod[[6]] <- metagen(
  coef, SE, data=aux,
  studlab=paste(authoryear),
  comb.fixed = FALSE,
  comb.random = TRUE,
  method.tau = "REML",
  hakn = TRUE,
  prediction=TRUE,
  sm="SMD"
  )

# Pooling effects analysis -- PCTGDP x K
aux <- dat %>%
  filter(indepvar2 == 'K',
         depvar2 == 'PCTGDP')

mod[[7]] <- metagen(coef, SE, data=aux,
          studlab=paste(authoryear),
          comb.fixed = FALSE,
          comb.random = TRUE,
          method.tau = "REML",
          hakn = TRUE,
          prediction=TRUE,
          sm="SMD")

# Full model - Pooling effects analysis -- PCTGDP x N
aux <- fulldat %>%
  filter(indepvar2 == 'N',
         depvar2 == 'PCTGDP')

mod[[8]] <- metagen(coef, SE, data = aux,
          studlab = paste(authoryear),
          comb.fixed = FALSE,
          comb.random = TRUE,
          method.tau = "REML",
          hakn = TRUE,
          prediction=TRUE,
          sm = "SMD")

# Full data - Pooling effects analysis -- ExpPC x K
aux <- fulldat %>%
  filter(indepvar2 == 'K',
         depvar2 == 'ExpPC')

mod[[9]] <- metagen(coef, SE, data=aux,
          studlab=paste(authoryear),
          comb.fixed = FALSE,
          comb.random = TRUE,
          method.tau = "REML",
          hakn = TRUE,
          prediction=TRUE,
          sm="SMD")
```

We then proceed to the meta-analysis. We matched the house size variables with
our measures of government spending and created a theoretical $3 \times 3$
matrix. Out of the 9 possible variable combinations, we found only 7 in the
article pool. Our sample does not contain any papers that analyse the
relationships between the log lower chamber size and public expenditure per
capita, or between upper chamber size and the logarithm of public expenditure
per capita. We calculated the effect sizes by drawing the coefficients and
standard errors from the articles and computing Hedges' $g$, which is a widely
used measure of standardised difference between means. When standard errors
were not directly available in the texts, we used t-statistics to retrieve
them.

Figure \ref{fig:plots} shows the forest plots for our restricted sample, which
includes the `r dim(dat)[1]` main coefficients of the `r length(unique(dat$id))`
selected papers.\footnote{Please refer to Sections H and I in the Supplementary
Material for full results regarding both samples.} On the left side of the
plots are the names of the study authors and paper publication year. For
unpublished studies, we included the first year the paper was available online.
The bars in the middle show the reported effect sizes and the vertical lines
indicate their average, weighted by standard errors. The length of the lines
represent the precision of the estimates. The red line at the bottom of the
figures displays the overall effects plus their respective confidence
intervals.

\begin{landscape}
\begin{figure}[!ht]
\centering
\caption{Forest plots of the relationship between legislature size and government spending (reduced sample)}
\vspace{0.3cm}
\includegraphics[width=25cm,height=17cm]{../graphs/graph1.pdf}
\label{fig:plots}
\end{figure}
\end{landscape}
